---
title: Molecular Docking Simulation
subtitle: Ayush Noori
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
# knitr::opts_chunk$set(eval = FALSE)
```

# Dependencies

Load requisite packages and define directories. Note that this script may also use my personal utilities package `brainstorm`, which can be downloaded via `devtools::install_github("ayushnoori/brainstorm")`.

```{r load-packages, message=FALSE, warning=FALSE}
# data manipulation
library(data.table)
library(purrr)
library(magrittr)

# relative file paths
base_dir = "../GEM 2024/generations_iclr_post" # edit path
library(here)

# data visualization
library(ggplot2)

# utility functions
library(brainstorm)

# cheminformatics tools
# remotes::install_github('jasenfinch/cheminf')
# library(cheminf)
# remotes::install_github("schymane/RChemMass")
# library(RChemMass)
```

Read files.

```{r read-files}
# read molecular analysis
mcts_mols = fread(here(base_dir, "mcts_chemprop_rdkit", "molecules.csv"))
random_mols = fread(here(base_dir, "random", "molecules_random.csv"))

# print quantiles
quantile(mcts_mols$chemprop_rdkit_ensemble_preds)
quantile(random_mols$chemprop_rdkit_ensemble_preds)
```

Get top 5 molecules and bottom 5 molecules by average DRD2 score.

```{r top-d2r}
# sort by d2r
mcts_mols = mcts_mols %>%
  .[, d2r_sum := chemprop_rdkit_model_drd2_model_0_preds + chemprop_rdkit_model_drd2_model_1_preds] %>%
  .[order(-d2r_sum)]

random_mols = random_mols %>%
  .[, d2r_sum := chemprop_rdkit_model_drd2_model_0_preds + chemprop_rdkit_model_drd2_model_1_preds] %>%
  .[order(-d2r_sum)]

# get top molecules
top_5 = mcts_mols[1:5, smiles]
bottom_5 = mcts_mols[(nrow(mcts_mols)-4):nrow(mcts_mols), smiles]

# docking results from SeamDock using AutoDock Vina
top_5_binding_affinity = c(-11, -11, -11.6, -11.9, -12)
bottom_5_binding_affinity = c(-6, -6, -7.4, -7.1, -5.3)
```

SMILES structure of risperidone retrieved from DrugBank [here](https://go.drugbank.com/drugs/DB00734). Create plot of molecular docking simulation results.

```{r plot-sim}
# create docking results
sim_res = data.table(
  Score = c(-11.4,top_5_binding_affinity, bottom_5_binding_affinity),
  Label = c("Risperidone", rep("High Scoring", 5), rep("Low Scoring", 5))
) %>%
  .[, Label := factor(Label, levels = c("Risperidone", "High Scoring", "Low Scoring"))]

# compute mean, SD, and SE by Label
summary_stats = sim_res[, .(
  Mean = mean(Score),
  SD = sd(Score),
  SE = sd(Score) / sqrt(.N)  # Standard Error
), by = Label]

# create plot
p = ggplot(summary_stats, aes(x = Label, y = Mean)) +
  geom_bar(stat = "identity", aes(fill = Label), color = "black") +
  scale_fill_manual(values = c("#DDD5E9", "#C7E5F8", "#FFCFD1")) +
  # scale_fill_manual(values = c("#7E5DAC", "#269BE3", "#FF474E")) +
  geom_point(data = sim_res, aes(y = Score), size = 3, shape = 21, color = "black", fill = "white") +
  geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE), width = 0.2, color = "black", linewidth = 0.8) +
  scale_y_reverse(expand = expansion(c(0, 0.05))) + 
  labs(x = "Category", y = "Binding Affinity (kcal/mol)") +
  theme_bw() +
  theme(
    axis.title.y = element_text(face = "bold", size = 12),
    axis.title.x = element_blank(),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12, face = "bold", color = "black"),
    legend.position = "none"
  )

# save plot
ggsave(here(base_dir, "molecular_simulation.png"), p, dpi = 900, height = 6, width = 5)
```

Read other statistics.

```{r read-stats}
mcts_dir = here(base_dir, "mcts_chemprop_rdkit", "chemprop_rdkit_ensemble_preds")

# read Tversky
b3db_tversky = fread(here(mcts_dir, "b3db_classification_preprocessed_tversky_similarity.csv"))
# nrow(b3db_tversky[max_tversky_similarity < 0.9])
excape_tversky = fread(here(mcts_dir, "drd2_binding_data_tversky_similarity.csv"))

# read Tanimoto
tanimoto = fread(here(mcts_dir, "internal_tanimoto_similarity.csv"))
# nrow(tanimoto[max_tanimoto_similarity < 0.8])
```
